Constructing a universe for the setoid model

### 2020-05-19 paper.tex
WIP

To provide such a universe we need to define a setoid $\setoidU : \Setoid$ and a
family of setoids $\El : \setoidU \to \Setoid$, with codes for basic type
formers like $\Pi$-types and booleans. This can be defined as the following
inductive-recursive type. \todo{expand definition to include more details}
%
\begin{align*}
  & \textsf{data} \ \setoidU : \Set_1 \\
  & \_\sim_\setoidU\_ : \setoidU \to \setoidU \to \Prop_1 \\
  & \El : \setoidU \to \Set \\
  & \_\vdash\_\sim_\El\_ : \{a\ a' : \setoidU\} \to a \sim_\setoidU a' \to \El\ a \to \El\ a' \to \Prop
\end{align*}

@@

As already pointed out, however, we would like to avoid having to assume
inductive-recursive types, of more generally induction-recursion, in the
metatheory.

Our result is that a modified form of this translation also works for
the more complex inductive-recursive type we need to model the universe of
setoids. 
%
In particular, in addition to $\ad{in-U}$ for defining $\U$ as before, we also
introduce a family $\ad{in-U∼}$ of binary relations between types in the
universe, from which we then define $\_\sim_{\setoidU}\_$.
%
\begin{align*}
  & \textsf{data}\ \ad{in-U} : \Set \to \Set_1 \\
  & \textsf{data}\ \ad{in-U∼} : \{A\ A' : \Set\} \to \ad{in-U}\ A \to \ad{in-U}\ A' \to (A \to A' \to \Prop) \to \Set_1
\end{align*}

>>>>

\subsection{A first definition of a setoid universe}

To provide such a universe we need to define a setoid $\setoidU : \Setoid$ and a
family of setoids $\El : \setoidU \to \Setoid$, with codes for basic type
formers like $\Pi$-types and booleans. That is, we need the following data
%
\begin{align*}
  & \textsf{data} \ \setoidU : \Set_1 \\
  & \_\sim_\setoidU\_ : \setoidU \to \setoidU \to \Prop_1 \\
  & \El : \setoidU \to \Set \\
  & \_\vdash\_\sim_\El\_ : \{a\ a' : \setoidU\} \to a \sim_\setoidU a' \to \El\ a \to \El\ a' \to \Prop
\end{align*}

@@

\todo{Acknowledge Simon Boulier for the IIR universe construction}
\todo{QIIT II LICS paper. this part is already in the HIIT paper}

We given an inductive definition of $\setoidU, ...$, mutually with a recursive
definition of $\El, \_\vdash\_\sim_\El\_, \sim_\setoidU, \textsf{refl}_\setoidU, ...$.

\begin{align*}
  & \textsf{data}\ \setoidU : \Set_1 \\
  & \quad \textsf{bool} : \setoidU \\
  %% Πs : (a : U i)(b : El a → U i)
     %% (refb : {x x' : El a}(x~ : refU a ⊢ x ~El x') → b x ~U b x') → U i
  & \quad \textsf{pi}
  :  (A : \setoidU) (B : \El\ A \to \setoidU)\\
  & \ \qquad \to (\{x\ x' : \El\ A\} \to \textsf{refl}_\setoidU\ A \vdash x \sim_\El x' \to B\ x \sim_\setoidU B\ x')
  \to \setoidU
\end{align*}

\[
\textsf{refl}_\setoidU : ?
\]

Note that in the definition of $\setoidU$, we crucially require that the family
$B : \El\ A \to \setoidU$ be a setoid morphism, hence that it respects the
corresponding setoid equalities.

This definition of the setoid universe relies on both induction-recursion and
induction-induction. Although appily accepted by Agda, the meaning of this
combination is not really understood. The goal of the next sections is therefore
to find an alternative definition of the setoid universe that relies on
principles that are better understood. Not only that, we would like to reduce
the number of extensions to the metatheory that are needed to be able to give
such definition, so to stay as close as possible to our initial idea of reducing
SeTT to a small core.  \todo{reword}

The next section explains our main result, which is that a modified form of the
translation of IR to IF also works for the setoid universe defined above. This
type is more complex that the basic IR types we have seen before, in that it
uses a combination of IR and II, and includes the universe of strict
propositions.

The reduction of the IIR universe to induction works, conceptually, as before.
%
However, in addition to $\ad{in-U}$ for defining $\U$, we also introduce a
family $\ad{in-U∼}$ of binary relations between types in the universe, from
which we then define $\_\sim_{\setoidU}\_$.
%
\begin{align*}
  & \textsf{data}\ \ad{in-U} : \Set \to \Set_1 \\
  & \textsf{data}\ \ad{in-U∼} : \{A\ A' : \Set\} \to \ad{in-U}\ A \to \ad{in-U}\ A' \to (A \to A' \to \Prop) \to \Set_1
\end{align*}
